---
title: "SS3 Webinars"
format: html
editor: 
  markdown: 
    wrap: 72
---

We have every other month webinars on SS3 features and topics. The first part of 
these webinars are more of a presentation on the feature or topic and the second
part is a Q&A session. Subscribe to the [SS3 Webinar calendar](https://calendar.google.com/calendar/u/0?cid=Y18zNWM2MWQ5Yjg1MGY1YjkwOGNiY2RjNjIwNmZlYzVkNDgxMmQ2ZTU0NDg3YjE0ZjM1ZGMzYmY0M2E3MGQyMmYxQGdyb3VwLmNhbGVuZGFyLmdvb2dsZS5jb20).

To suggest future webinar topics, please post to the [SS3 User Forum Google Group](https://groups.google.com/g/ss3-forum).

## Impact of time-varying biology on reference points in SS3
Date is TBD.
```{r, echo = FALSE, out.width = "90%", out.height= "400px", fig.alt="pdf of presentation from Rick Methot on the impact of time-varying biology on reference points in SS3"}
knitr::include_graphics(normalizePath("../qmds/presentations/TVBiology_SR_RefPoints.pdf"))
```

## January 16, 2025 - Richards curve with Arni Magnusson

```{r, echo = FALSE, out.width = "90%", out.height= "400px", fig.alt="pdf of presentation from Arni Magnusson on the Richards Curve"}
knitr::include_graphics(normalizePath("../qmds/presentations/Richards_Curve_Webinar.pdf"))
```

Please also see the [Richards Growth Curve vignette](qmds/richards_growth_curve.qmd).

## November 19, 2024 - Interacting with SS3 repos on GitHub
```{r, echo = FALSE, out.width = "90%", out.height= "400px", fig.alt="pdf of presentation on interacting with SS3 repos on GitHub"}
knitr::include_graphics(normalizePath("../qmds/presentations/SS3_GitHub_Workshop.pdf"))
```

## October 15, 2024 - Recruitment with time-varying biology
**Materials will appear when approved for public consumption.**

## May 16, 2024 - Fishing Mortality ($F$)

```{r, echo = FALSE}
path_files <- list.files(path = file.path(getwd(), "example_files", "demo_constant_timeseries"), full.names = TRUE)

downloadthis::download_file(
  path = path_files,
  output_name = "f_mortality",
  button_label = "Download F Mortality Example input files",
  button_type = "default",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```
  
<br/><br/>

```{r, echo = FALSE, out.width = "90%", out.height= "400px", fig.alt="pdf of Fishing Mortality presentation"}
knitr::include_graphics(normalizePath("../qmds/presentations/SS3_Fishing_Mortality.pdf"))
```
